Open source landscape ecology tools \thanks{Grants or other notes about the article that should go on the front page should be placed here. General acknowledgments should be placed at the end of the article.}

### 2020-10-10 R/survey_stats.Rmd
Starting to work on results of survey

new_names = c("time", "position", "how_often", "expertise",
              "major_topic", "important_tasks", "data_model",
              "r_packages", "own_package", "own_package_location",
              "missing_methods", "usefulness", "comments")

@@

Number of answers - `r nrow(x)`.

@@

```{r}
# how_often
ggplot(x, aes(y = fct_infreq(how_often))) +
    geom_bar() +
    labs(x = "Count", y = NULL, title = old_names[3])
```

>>>>

new_names = c("time", "position", "usage_freq", "expertise", "major_topic",
              "important_tasks", "data_model", "r_packages", "own_package",
              "own_package_location", "missing_methods", "usefulness", "comments")

@@

Number of answers: `r nrow(x)`.

@@

```{r}
# how_often
ggplot(x, aes(y = fct_infreq(usage_freq))) +
    geom_bar() +
    labs(x = "Count", y = NULL, title = old_names[3])
```

### 2020-10-10 paper/paper.Rmd
Starting to work on results of survey

@@

## Spatial regression
<!-- MH: Do we really want to include this? Not really a landscape ecology topic but rather (geo) spatial stats I would argue -->
<!-- JN: we could wait with this section to after receiving the results of the questionnaire -->

@@

>>>>

```{r setup R, include = FALSE}
# load packages
library(dplyr)
library(forcats)
library(ggplot2)
library(googlesheets4)
library(stringr)
library(tidyr)

answers <- read_sheet("https://docs.google.com/spreadsheets/d/1mV_wZWb4desjdzUWHALIB-uFRCauSTOGSvENsnkiAhI/edit#gid=1189808333")

old_names <- names(answers)

new_names <- c("time", "position", "usage_freq", "expertise", "major_topic",
               "important_tasks", "data_model", "r_packages", "own_package",
               "own_package_location", "missing_methods", "usefulness", "comments")

names(answers) <-  new_names
```

@@

@@

# Survey of R usage by landscape ecology community

```{r participation, include = FALSE}
digits <- 1

# number of people that took part
n_people <- nrow(answers)

# current position
current_position <- dplyr::group_by(answers, position) %>% 
  dplyr::summarise(n_rel = n() / n_people * 100) %>% 
  dplyr::mutate(n_rel = round(n_rel, digits = digits)) %>% 
  dplyr::arrange(-n_rel)

# how often do you use R?
usage_freq <- dplyr::group_by(answers, usage_freq) %>% 
  dplyr::summarise(n_rel = n() / n_people * 100) %>% 
  dplyr::mutate(n_rel = round(n_rel, digits = digits)) %>% 
  dplyr::arrange(-n_rel)

# expertise
expertise <- dplyr::group_by(answers, expertise) %>% 
  dplyr::summarise(n_rel = n() / n_people * 100) %>% 
  dplyr::mutate(n_rel = round(n_rel, digits = digits)) %>% 
  dplyr::arrange(-n_rel)
```

To better understand how *R* is used by the landscape ecology community, we designed a short survey and used mailing lists and social media to reach the landscape ecology community. 
In total, the survey was answered by `r n_people` people, of which the the majority were either "PhD students" (`r current_position[[1, 2]]`%), followed by "Post-Docs" (`r current_position[[2, 2]]`%) and "Professors" (`r current_position[[3, 2]]`%). Other less frequent answers were "data scientists", "government employees" or "bachelor's and master's degree students". 
Most people used R either "daily" (`r usage_freq[[1, 2]]`%) or a "few times a week" (`r usage_freq[[2, 2]]`%) and and about half of all participants described themselves as "advanced" user (`r expertise[[1, 2]]`%), while still `r expertise[[2, 2]]`% described themselves as "intermediate" users.

```{r data_model_packages, include = FALSE}
# data model
data_model <- dplyr::group_by(answers, data_model) %>% 
  dplyr::summarise(n_rel = n() / n_people * 100) %>% 
  dplyr::mutate(n_rel = round(n_rel, digits = digits)) %>% 
  dplyr::arrange(-n_rel)

# split answers r packages
r_packages <- unlist(str_split(answers$r_packages, pattern = ","))
r_packages <- unlist(str_split(r_packages, pattern = ";"))
r_packages <- unlist(str_split(r_packages, pattern = "\\*"))
r_packages <- str_trim(r_packages)

# clean some special cases
r_packages[str_detect(r_packages, pattern = "landscape metrics")] <- "landscapemetrics"
r_packages[str_detect(r_packages, pattern = "spectre*")] <- "spectre"
r_packages[str_detect(r_packages, pattern = "map tools")] <- "maptools"
r_packages[str_detect(r_packages, pattern = "gdal and many more")] <- "rgdal"
r_packages[str_detect(r_packages, pattern = "I would like to know")] <- "NA"
r_packages[str_detect(r_packages, pattern = "no")] <- "NA"
r_packages[str_detect(r_packages, pattern = "currently none")] <- "NA"
r_packages[!str_detect(r_packages, pattern = "")] <- "NA"

# split last packages
r_packages <- unlist(str_split(r_packages, pattern = " "))

# remove NA answers
r_packages <- r_packages[!str_detect(r_packages, pattern = "NA")]

# make sure all are lower case
r_packages <- str_to_lower(r_packages)

# create data.frame
r_packages_df <- tibble(r_package = r_packages) %>%
  dplyr::group_by(r_package) %>% 
  dplyr::summarise(n = dplyr::n()) %>% 
  dplyr::mutate(n_rel = n / sum(n) * 100, 
                r_package_reclass = dplyr::case_when(n >= 5 ~ r_package, 
                                                     n < 5 ~ "others")) %>%
  dplyr::group_by(r_package_reclass) %>% 
  dplyr::summarise(n_rel = sum(n_rel)) %>% 
  dplyr::mutate(label = paste0(round(n_rel, digits = 1), "%"), 
                col = dplyr::case_when(r_package_reclass == "others" ~ 0, 
                                       r_package_reclass != "others" ~ 1), 
                col = factor(col)) %>% 
  dplyr::arrange(-n_rel)

# change ordering
new_levels <- c("others", rev(r_packages_df$r_package_reclass[-1]))

# reorder levels
r_packages_df$r_package_reclass <- factor(r_packages_df$r_package_reclass, 
                                          levels = new_levels)
```
Interestingly, most people used the raster data model (`r data_model[[1, 2]]`%) in comparison to the vector data model (`r data_model[[2, 2]]`%).

```{r used_package, echo = FALSE, fig.align = "center"}
ggplot(data = r_packages_df) +
  geom_bar(aes(y = r_package_reclass, x = n_rel, fill = col), stat = "identity") + 
  geom_text(aes(y = r_package_reclass, x = n_rel, label = label, col = col), 
            nudge_x = 2.5) + 
  theme_classic() + 
  scale_x_continuous(breaks = seq(0, 45, 5), limits = c(0, 45)) + 
  scale_fill_manual(values = c("grey50", "black")) +
  scale_colour_manual(values = c("grey50", "black")) +
  guides(fill = FALSE, colour = FALSE) +
  labs(x = "Relative value [%]", y = expression(paste(italic(R), " package")))
```

### 2020-10-10 paper/paper.tex
Starting to work on results of survey

\hypertarget{spatial-regression}{%
\subsection{Spatial regression}\label{spatial-regression}}

@@

>>>>

@@

\hypertarget{survey-of-r-usage-by-landscape-ecology-community}{%
\section{Survey of R usage by landscape ecology community}\label{survey-of-r-usage-by-landscape-ecology-community}}

To better understand how \emph{R} is used by the landscape ecology community, we designed a short survey and used mailing lists and social media to reach the landscape ecology community.
In total, the survey was answered by 94 people, of which the the majority were either ``PhD students'' (33\%), followed by ``Post-Docs'' (26.6\%) and ``Professors'' (12.8\%). Other less frequent answers were ``data scientists'', ``government employees'' or ``bachelor's and master's degree students''.
Most people used R either ``daily'' (58.5\%) or a ``few times a week'' (34\%) and and about half of all participants described themselves as ``advanced'' user (48.9\%), while still 39.4\% described themselves as ``intermediate'' users.

Interestingly, most people used the raster data model (70.2\%) in comparison to the vector data model (29.8\%).

\begin{center}\includegraphics{paper_files/figure-latex/used_package-1} \end{center}
